Hilbert space approximate Bayesian Gaussian processes using Stan

### 2020-03-26 Paper/manuscript_v2/manuscript_07_01.tex
More editing and to dos for sec 4

Next, we will focus on analyzing the interaction effects between these $m$ and $c$ on the performance of the approximation. The lengthscale and marginal variance of the kernel will no longer be fixed but rather estimated in both regular GP and HSGP models. Figure \ref{fig3_Post_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$. Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP model. Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models. Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$. Further, the choice of $m = 10$ and $c = 1.5$ could still be an accurate enough choice. We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig3_Post_part1}. From these results, some general conclusions may be drawn:

\begin{itemize}
\item As $c$ increases, $m$ has to increase as well (and vice versa).
\item There exists a minimum $c$ below which a close approximation will never be achieved regardless of $m$.
\end{itemize}

@@

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depicts how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension. More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying equation (\ref{eq_diff_covs}). Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function. We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function:
%
\begin{equation*}
 \frac{\varepsilon}{\int k(\tau) \mathrm{d}\tau} < 0.01.
\end{equation*}

@@

As stated above, Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used). If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, $\hat{\ell}$ does not exceed recommended minimum lengthscale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased. We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
As mentioned earlier, $c$ cannot be decreased as much as desired because it is restricted by the lengthscale, so increasing $m$ may usually the preferred approach.

>>>>

Next, we analyze how the interaction effects between $m$ and $c$ affects the quality of the approximation. The lengthscale and marginal variance of the covariance function will no longer be fixed but instead estimated in both regular GP and HSGP models. \todo{Specify how the hyperparameters are handled} Figure \ref{fig3_Post_part1} shows the posterior predictive mean of the function and the covariance function obtained after fitting the data for varying $m$ and $c$. Figure \ref{fig4_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models computed against the regular GP model. Figure \ref{fig5_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models. Looking at the RMSEs in Figure \ref{fig4_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$. Further, the less conservative choice of $m = 10$ and $c = 1.5$  could also produce a sufficiently accurately approximation depending on the application. We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig3_Post_part1}. From these results, some general conclusions may be drawn:

\begin{itemize}
\item As $c$ increases, $m$ has to increase as well (and vice versa). This is consistent with the expression for the eigenvalues in eq. \eqref{eq_eigenvalue}, where $L$ appears in the denominator	.
\item There exists a minimum $c$ below which a accurate approximation will never be achieved regardless of the number of basis functions $m$.
\end{itemize}

@@

Additionally, there is a clear relation between the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} depict how these three factors interact with each other in relation to a close approximation of the HSGP model, in the cases of a GP with square exponential covariance function and Mat\'ern($\nu$=3/2) covariance function, respectively, and a single input dimension. More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a accurate approximation in the sense of satisfying equation (\ref{eq_diff_covs}). Similarly for Figure \ref{fig6_lscale_vs_J_vs_c_Matern} in the case of a Mat\'ern($\nu$=3/2) covariance function. We considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in equation (\ref{eq_diff_covs}), is below 1$\%$ of the density of the exact covariance function
\todo{1) I don't fully understand the statement in the sentence about. Which density are you referring to? 2) I also think the is a problem with the equation below. The integral $\int k(\tau) \mathrm{d}\tau$ is just a constant (for fixed hyperparameters). Furthermore, we need to use different symbols to represent the exact covariance function and the approximate covariance function, you could use $k$ and $\tilde{k}$ for example. }
%
\begin{equation*}
 \frac{\varepsilon}{\int k(\tau) \mathrm{d}\tau} < 0.01.
\end{equation*}

@@

As stated above, Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern} provide the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the function, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig6_lscale_vs_J_vs_c} or \ref{fig6_lscale_vs_J_vs_c_Matern} (depending on which kernel is used). If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, $\hat{\ell}$ does not exceed recommended minimum lengthscale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased. We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale, repeat the process while increasing $m$ or decreasing $c$.
As mentioned earlier, $c$ cannot be decreased as much as desired because it is restricted by the lengthscale, so increasing $m$ may usually the preferred approach.
